An OpenCL-based tridiagonal solver for evaluation of compact finite differences on parallel architectures

### 2015-09-20 report.tex
add tesla block diagram; update report

\section{GPU architecture and parallelization strategies}

@@

    \subsection{Results for a single node}

    \begin{table}
    \centering
    \begin{tabular}{|l|r|r|r|r|}
    \hline
    Size    &      CPU (1 core) &  CPU (8 cores) & GPU (without memory transfers)  & GPU (with memory transfers) \\
    \hline
    32      &     0.0005        &   0.00029     &   0.00128 &   0.0014   \\
    64      &     0.0038        &   0.0008      &   0.0018  &   0.00385  \\
    128     &     0.042         &   0.0075      &   0.0073  &   0.02014  \\
    256     &     0.34          &   0.062       &   0.057   &   0.148    \\
    512     &     2.45          &   0.47        &   0.70    &   1.401    \\
    \hline
    \end{tabular}
    \caption{Performance of solvers on a single GPU (Tesla k20m)}
    \label{table:res1}
    \end{table}

    The results for a single node bring up some issues in performance evaluation.
    First, it is clear that the major cost for the GPU is not in the computational
    kernels, but rather in transferring data between host (CPU) and device (GPU) buffers.
    In the context of a larger piece of software,
    if only the computation of the compact finite difference was offloaded
    to the GPU, then considering this cost is absolutely essential.
    However, if the rest of the problem is also written to be solved on the GPU,
    then there is no need for transfers to and from the host at each
    time step.

>>>>

\section{GPU architecture and OpenCL programming platform}

    \begin{figure}[h]
    \begin{center}
    \includegraphics[height=200pt]{img/tesla-block-diagram.jpg}
    \end{center}
    \caption{Distribution of a 2-D domain among processes}
    \label{fig:da-domain}
    \end{figure}

    A graphics processing unit (GPU) consists of several (thousand)
    compute cores and is designed for high-throughput computations.

    NVIDIA cards support extensions such as CUDA and OpenCL for
    langauages like C, C++, Fortran and Python.
    Applications written using these extensions are launched
    with thousands of lightweight threads,
    each operating on a small portion the problem domain (typically a single grid point).

    Each GPU has several so-called streaming multiprocessors (SMs)---the Tesla K20M
    has 13 SMs---each of which has its own control units, registers, caches, etc.
    Instructions execute simultaneously on an SM.

    All SMs (and therefore all threads) have access to the GPU's ``global'' memory.
    Access to global memory is efficient when concurrent threads access neighbouring memory locations
    (known as \emph{coalesced memory access}).

    All threads in an SM also have access to a common fast, on-chip, \emph{shared memory}.
    Access to shared memory, in general, is much faster than access to global memory,
    but the amount of shared memory per SM is limited, typically a few dozen KB.
    Certain memory access patterns by concurrent threads to shared memory may result in
    serialization of the memory accesses, known as bank conflicts. These are detailed [here].

    The OpenCL programming platform

\section{Parallelization strategies}

@@

    \subsection{Results for a single node}
    The results for a single node bring up some issues in performance evaluation.
    First, it is clear that the major cost for the GPU is not in the computational
    kernels, but rather in transferring data between host (CPU) and device (GPU) buffers.
    In the context of a larger piece of software,
    if only the computation of the compact finite difference was offloaded
    to the GPU, then considering this cost is absolutely essential.
    However, if the rest of the problem is also written to be solved on the GPU,
    then there is no need for transfers to and from the host at each
    time step.
